---
title: "Desafio Solarview"
author: "Semar Augusto"
output:
  html_document:
    df_print: paged
---

```{r, warning=FALSE, include=F}
library(tidyverse)
library(rgdal)
library(reshape2)
library(tseries)
library(forecast)
library(RColorBrewer)  # for brewer.pal(...)
```


# Data loading

```{r}
states <- readOGR("UFEBRASIL.shp", encoding = "UTF-8")
df <- read.csv("radiacao_sql.csv", encoding = "UTF-8")
```

# Initial Description

```{r}
summary(df)

means <- as.data.frame((apply(df[,4:16], 2, mean)))
colnames(means) <- "Mean"
means$names <- factor(c(month.name, "Annual"), levels=c(month.name, "Annual"))
ggplot(means, aes(x = names, y = Mean)) + geom_bar(stat='identity')
```

I would expect that there would be a big difference in the mean and median whilst comparing winter and summer time. It seems not to be the case. However, Brazil is a big country and has a big part of its land area in the Equador. Let's check if in a smaller region further away from Equador will express this difference.

```{r}
sc <- subset(df, STATE == "SANTA CATARINA")
summary(sc)
means <- as.data.frame((apply(sc[,4:16], 2, mean)))
colnames(means) <- "Mean"
means$names <- factor(c(month.name, "Annual"), levels=c(month.name, "Annual"))
ggplot(means, aes(x = names, y = Mean)) + geom_bar(stat='identity')
rm(sc)
```

As expected, a region further away from the Equador has bigger differences between summer and winter.

We'll check now if by taking a a state closer to Equador will have simillar difference. That would mean that the fact that there is little difference overall in Brazil is due to a sum of variabilities

```{r}
am <- subset(df, STATE == "AMAZONAS")
summary(am)
means <- as.data.frame((apply(am[,4:16], 2, mean)))
colnames(means) <- "Mean"
means$names <- factor(c(month.name, "Annual"), levels=c(month.name, "Annual"))
ggplot(means, aes(x = names, y = Mean)) + geom_bar(stat='identity')
rm(am)
```

As expected there is actually little variability in a state close to Equador.

```{r}
df1 <- df
df1 <- aggregate(df1[,16], by=list(df1$YEAR), mean)
colnames(df1)[1] <- "YEAR"
ggplot(df1, aes(x=YEAR, y=x)) + geom_line()
```

Solar radiation seems to have pumped up by 1995 and hasn't come down since.

# Time Series Analysis

I'll transform the data frame in order to each month be a line. That will make it easier on the time series algorithms.
```{r}
df1 <- select(df, -"ANN")
df1 <- subset(df, STATE == "MINAS GERAIS")
df1 <- aggregate(df1[4:15], by=list(df1$YEAR, df1$STATE), mean)
colnames(df1)[1] <- "YEAR"
colnames(df1)[2] <- "STATE"
df1 <- melt(df1, id.vars=c("YEAR", "STATE"))

tsdf <- ts(df1[, c("value")])
tsdf <- ts(as.vector(tsdf), start = c(1984,1), frequency = 12)

components.ts <-  decompose(tsdf)
plot(components.ts$x)
```

```{r}
plot(components.ts$random)
```

```{r}
decomp = stl(tsdf, s.window="periodic")
plot(decomp) 
```

# Prediction

We'd like to predict what the next two years of radiation will be like. In order to do so, a ARIMA model is very common.

However, it requires stationary data. Let's check if that is the case in our dataset.
```{r}
adf.test(tsdf, alternative = "stationary")
```

The formal ADF test does not reject null hypothesis, which means the dataset used is most likely a non-stationary one.

```{r}
Acf(tsdf, main='ACF for Differenced Series')
Pacf(tsdf, main='PACF for Differenced Series') 
```


We'll use seasadj function to try to remove seasonality
```{r}
deseasonal_cnt <- seasadj(decomp)
count_d1 = diff(deseasonal_cnt, differences = 1)
plot(count_d1)
adf.test(count_d1, alternative = "stationary")
```

```{r}
Acf(count_d1, main='ACF for Differenced Series')
Pacf(count_d1, main='PACF for Differenced Series') 
```

Got our stationary dataset. Let's proceed with the prediction process

# Predicting next two years

```{r}
fit_ts <- auto.arima(deseasonal_cnt, seasonal=FALSE)
fit_ts2 <- ets(tsdf)
fit_ts
fit_ts2
```

Arima model has better AIC, AICc and BIC, we'll proceed with it.

```{r}
tsdisplay(residuals(fit_ts), lag.max=45, main='(1,1,1) Model Residuals') 
```

```{r}
fcast <- forecast(fit_ts, h=24)
plot(fcast)
```

# Spatial Description

```{r}
df1 <- subset(df, YEAR == 2005)
df3 <- df1[complete.cases(df1), c("LON", "LAT", "DEC")]
ggplot(df3[complete.cases(df3),], aes(x=LON, y=LAT, color=DEC)) + 
    ggtitle("Solar Radiation at December 2005") + 
    geom_point()+
    geom_path(data=states,aes(x=long, y=lat,group=group), colour="grey50")+
    scale_fill_gradientn(colours=rev(brewer.pal(7,"Spectral")))+
    coord_fixed()
```

```{r}
df4 <- df1[complete.cases(df1), c("LON", "LAT", "JUN")]
ggplot(df4[complete.cases(df4),], aes(x=LON, y=LAT, color=JUN)) + 
    ggtitle("Solar Radiation at June 2005") + 
    geom_point()+
    geom_path(data=states,aes(x=long, y=lat,group=group), colour="grey50")+
    scale_fill_gradientn(colours=rev(brewer.pal(7,"Spectral")))+
    coord_fixed()
```



